---
title: "data_manipulation"
output: html_document
date: "2023-09-21"
---

```{r}
library(tidyverse)
```

Import FAS litters and pups

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## select

`select` is used to select columns

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, gd0_weight, group)

select(litters_df, group, starts_with("pups"))

select(litters_df, -litter_number)

select(litters_df, -starts_with("GD"))

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())
```

```{r}
rename(litters_df, litter_id = litter_number)

relocate(litters_df, litter_number)
```

# learning assessment 1: in the pups dataset, select the columns containing litter_number, sex, pd_ears

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

`select` vs `pull`

```{r}
select(litters_df, group)
pull(litters_df, group)
```
## `filter`

get rid of rows using `filter()`

```{r}
filter(litters_df, group == "Con7")

filter(litters_df, group == "Mod8")

filter(litters_df, group != "Mod8")

filter(litters_df, gd0_weight > 20)

filter(litters_df, gd0_weight <= 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, (!group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20)
```

# learning assessment 2 - go back!!! 

```{r}





```

## `mutate` 

this is used to add or change variables to the dataframe (see by `view()` in the console)

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## to remove missing values or "NAs", `drop_na`: 

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```

## `arrange`


```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, group, gd0_weight)
```

## pipes

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(-starts_with("pups")) |>
  mutate(
    group = str_to_lower(group),
    wt_gain = gd18_weight - gd0_weight,
  ) |> 
  drop_na(wt_gain) |> 
  arrange(group, wt_gain)

```
To run the code for `litters_df`, run lines 131 - 139. To save the code into `litters_df`, run lines 130 - 139 (include `litters_df=`). View saved, new, `litters_df` in console by: `view(litters_df)` after running lines 130-139. 















